---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# gridbr: IBGE statistical grid access with R

<!-- badges: start -->
[![R-CMD-check](https://github.com/luisfelipebr/gridbr/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/luisfelipebr/gridbr/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

**The goal of 'gridbr' is to provide easy access to [IBGE Statistical Grid](https://mapasinterativos.ibge.gov.br/grade/default.html)** (download available [here](https://geoftp.ibge.gov.br/recortes_para_fins_estatisticos/grade_estatistica/censo_2010/)). It builds upon the development made in the project [IBGE Statistical Grid in Compact Representation](https://github.com/osm-codes/BR_IBGE) (short paper available [here](http://mtc-m16c.sid.inpe.br/ibi/8JMKD3MGPDW34P/45U7J5H)).

## Installation

You can install the development version of *gridbr* from this GitHub repository with `devtools`:

``` r
devtools::install_github("luisfelipebr/gridbr")
```

## Example

```{r setup}
library(gridbr)
library(sf)
```

**gridbr_download** is the main function available in the package. It builds the geometry of IBGE statistical grid and merge it with 2010 population census data. The user must provide the area of interest (aoi) as a geospatial [sf](https://r-spatial.github.io/sf/index.html) object.

In the following example, the Brazilian archipelago [Fernando de Noronha](https://en.wikipedia.org/wiki/Fernando_de_Noronha) is used as aoi. It can be download in one line with the package [geobr](https://ipeagit.github.io/geobr/) just by using its IBGE municipality code.

```{r}
library(geobr)
aoi <- read_municipality(2605459)
plot(st_geometry(aoi))
```

After loading the aoi, you can use the function `gridbr_download` to download the grid. Two arguments are mandatory: `input` and `cellsize`.

```{r}
aoi_grid <- gridbr_download(input = aoi, 
                            cellsize = "1KM")
head(aoi_grid)
plot(st_geometry(aoi_grid))
plot(st_geometry(aoi), add = TRUE, border = "red")
```

The IBGE statistical grid was made available in the following sizes: 500KM, 100KM, 50KM, 10KM, 5KM, 1KM and 200M. Using a different value will result in an error.

Retrieving population census data requires either an internet connection or the support package [gridbr.data](https://github.com/luisfelipebr/gridbr.data). If the user does not meet these requirements, the output will include only the cell's id and geometry, but not population census data. If the user meet the requirements and still don't want to download the data, they can specify it by setting the `census_data` parameter to FALSE.

```{r}
aoi_grid_1KM_id <- gridbr_download(aoi, 
                                   cellsize = "1KM", 
                                   census_data = FALSE)
head(aoi_grid_1KM_id)
```

The other optional parameter is related to the projection: by default, the output will have the same CRS as the input. If you want to keep the original grid CRS, with an equal area projection, you must set the `equal_area` parameter to TRUE.

```{r}
aoi_grid_1KM_id <- gridbr_download(aoi, 
                                   cellsize = "1KM", 
                                   equal_area = TRUE)
head(aoi_grid_1KM_id)
```
